### 2022-09-23 manuscript/manuscript.tex
some writing

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}
% \DeclareMathOperator{\var}{var}
% \DeclareMathOperator{\cov}{cov}

@@

>>>>

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}

@@

Signal processing

Genomes have inherent structure that impacts the linked selection map. There is
tremendous heterogeneity across species in the organization of likely selected
regions on the genetic and physical maps.

Within an organism, there are two key determinants of B maps: recombination and
gene density.

Initially, the relationship between recombination and diversity was regressed
with no regard to the spatial organization of the genome.

Certain regions are pinchpoints of selection: some combination of low
recombination and high gene density lead to very strong reductions in genetic
diversity. In these regions, coalescent dynamics change considerably; where as
under strong background selection, deleterious alleles would not fix, in these
regions the can fix. Additionally, the selective interference impacts positive
selection.

\section*{Theory}

We model the reduction in pairwise coalescence times along the genome due to
background selection by extending the work of \textcite{Santiago2016-mu}
(hereafter, the SC16 model). The SC16 model derives from a class of linked
selection \parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs} that
predict the reduction in pairwise coalescence time at a focal neutral site
using macro-state parameters such as fitness variation, rather than micro-state
parameters like the frequency trajectories of selected alleles (e.g.
\textcite{Maynard_Smith1974-zr}). We will briefly introduce the SC16 model's
approach and how we validated this model could be extended to whole-genome
inference.

Classic background selection theory models how coalescence times are reduced by
deleterious mutation-selection balance at neighboring sites. Building upon
older models of genetic load \parencite{Crow1970-wj,Kimura1966-bk}, these
models consider the equilibrium distribution of the \emph{number} of
deleterious mutations segregating on haplotypes
\parencite{Charlesworth1993-gb,Nordborg1996-nq}. Given an infinitely large
population, or sufficiently strong selection, the deleterious mutations in
these models cannot fix; consequently, any lineage containing

While the effect of selection on linked selection distorts neighboring
genealogies (XXX), skewing the site frequency spectrum (XXX) and altering
linkage disequilibria along the chromosome, we focus on the impact of selection
on pairwise coalescence rates. This simple summary is amenable to mathematical
and statistical inference, and is robust to difficulties in identifying rare
variants. The effect of background selection on pairwise coalescence rates is
expressed as the reduction ratio $B = \nicefrac{N_e}{N}$ where $N$ is the
drift-effective population size (i.e. including demographic processes) and
$N_e$ is the effective population size under background selection.

Under quantitative genetic models of linked selection, the reduction in
effective population size is determined by the product of fitness variation,
$V_f$, and an inflation factor $Q$. It is helpful to understand this by comparing
it to a model of $N_e$ without \emph{heritable} fitness variation but with
\emph{non-heritable} fitness variation. In this case,

\begin{align}
  N_e = \frac{4N}{2 + V_k},
\end{align}

where $V_k$ is the variance in offspring number across individuals. Under a
constant population size Wright--Fisher model of reproduction, $V_k \approx 2$,
and this simplifies to $N_e = N$. However, as first noted by
\textcite{Robertson1961-ho}, linked selection inflates the variance in
offspring number, as a neutral allele that becomes associated with a
high-fitness or low-fitness genetic background has an $1-r$ chance of remaining
associated with that background in the next generation, where $r$ is the chance
of recombination (or independent assortment, $r=\nicefrac{1}{2}$).

This persistence of this association inflates the effective variance in fitness
by a factory $Q^2$,

\begin{align}
  N_e = \frac{4N}{2 + V_k + 4 Q^2 V_f},
\end{align}

\begin{align}
  N_e = \frac{4N}{2 + V_k + 4 Q(r)^2 V_f},
\end{align}

\begin{align}
  N_e = N \exp\left(-Q^2 V_f\right)
\end{align}

In a region of $L$ basepairs under purifying selection, there is a net flux of
deleterious mutations in at rate $U = 2 \mu L$ per generation (where the factor
of two is because in diploids, $2L$ is the total number of basepairs per genome).

Typical models of background selection consider the equilibrium distribution of
the \emph{count} of deleterious mutations, rather than then level of fitness
variation implied by this count. If we assume fixed additive effects across
alleles and multiplicative fitness effects across loci, each copy of a
deleterious mutation an individual carries reduces their fitness by a factor of
$1-s$. The total reduction for an individual is

\begin{align}
  x = (1-s)^n
\end{align}

where $n$ is the number of deleterious alleles an individual carries. Under
mutation-selection balance, the population evolves to either a stationary
distribution under the strong BGS domain, or a traveling distribution in weak
selection case. Under some assumptions, equilibrium models of the strong domain
show that the distribution $f(n) \sim \text{Pois}(\nicefrac{U}{s})$
\parencite{Kimura1966-bk,Crow1970-wj} (CHECK) in the asexual case, and the
fixation rate of deleterious mutations per generation, or ratchet rate, is $R =
0$ since the probability of fixation, $p_F$ is zero. By contrast, as the
selection coefficient grows weaker, $s \to 0$, the probability of fixation $p_F
\to \nicefrac{1}{2N}$, and the ratchet approaches the neutral rate of $R =
\mu$. The weak selection domain is of importance because both because it
implies a certain likely irreversible rate of fitness reduction, and because it
could possibly confound population genetic inference which assumes dynamics
only under the strong model.

As $s \to 0$, the fitness distribution, put on a scale reflecting the fitness
relative at one moment in time, travels leftwards and changes shape. It has
been difficult to find dynamic equations for the rate of the ratchet, since
lower moments of the distribution, such as the variation, depend on higher
moments, such as the skew \parencite{Good2013-lp,Haigh1978-gt,Higgs1995-xc}.
This moment closure problem occurs regardless of whether one models the count
number of deleterious mutations (XXX) or the fitness distribution (XXX).

We can approximate the mapping between the distribution of deleterious
mutations in the population, and the fitness distribution.

Assuming $s$ is small, we can write express this in terms of the additive
genetic variance,

\begin{align}
  V(f) \approx s^2 \V(n) + \delta_\text{LD}.
\end{align}

The first term is the additive genic variance, which is

Under this model, the rate of the ratchet $R = U  - s V(n)$, which implies,

\begin{align}
  R = U - \nicefrac{V(f)}{s} \\
  V(f) = s(U - R)
\end{align}

\begin{align}

\end{align}

- We need to mention the two heterozygosities (S&C '16 says they diverge under
$N_e s < 1$ -- could explain anything we see?)

- A point about reconciling the two micro and macro-state theories of
population and quantitative genetics.

\subsection*{The Rate of the Ratchet}

\section*{Methods}

\subsection*{Data Preprocessing}

\subsection*{Human}

Site-level neutral and accessibility masks

Bin-level masks based on fraction of accessible bases in a window (7\%).

Bin-level tail filtering.
